---
title: |
  | Symanto Research for Text Analytics
  | API. Insights on texts using Symanto AI technology.
author: "Symanto"
date: "`r Sys.Date()`" 
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Symanto Research for Text Analytics. API. Insights on texts using Symanto AI technology.}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This document explains how to use the R package **symantoAPI** for doing text analytics. The methods available extract a bunch of psychological features from texts, by calling both the Symanto Application Programming Interface (Symanto API) and other APIs. The main code to call each funtion is shown below.

```{r setup, include=FALSE}
library(knitr)       # For knitting document and include_graphics.
library(kableExtra)  # Displays advanced tables.
knitr::opts_chunk$set(echo = TRUE)
```

The R package **symantoAPI** is loaded with the following command:

```{r packages, message=FALSE}
library(symantoAPI) 
```

The Symanto API is available at this link:

```{r api}
url_api <- "https://api.symanto.net"
```

Before using the Symanto API, a token key must be previously requested at [Symanto developer portal.](https://developers.symanto.net/signup)

```{r token request, eval=FALSE}
token <- "Request your API Key: https://developers.symanto.net/signup"
```

# Usage note

All the following methods can be run both with single texts and with a bunch of texts. Due to the internal procedure of the computational model developed, the methods available as **get_xxx()** (except for language detection) only allow sets of texts of the same language. In order to obtain the psychological features from a set of texts of different languages, the function **label_text()** must be used.

# Methods

The following toy text will be used for illustrative purposes:

```{r get personality comment}
text <- "Hello I love the service."
language <- "en"
```

Below is detailed every single method.

### Personality traits

The function **get_personality()** allows the user to obtain personality traits. These traits evaluate the content and writing style to determine how the author makes decisions. The model identifies whether an author is *emotional* (relationship-oriented, focusing on social values and empathy) or *rational* (objective and pragmatic, focusing on facts and logical deduction).

The input and output of the function are detailed next.

* **INPUT:**
  + URL to the API.
  + Text to analyze. A set of maximum 32 texts is allowed.
  + Language in which the text is written. Supported languages are "ar" (Arabic), "de" (German), "en" (English), "es" (Spanish), "fr" (French), "it" (Italian), "nl" (Dutch), "pt" (Portuguese), "ru" (Russian), "tr" (Turkish) and "zh" (Chinese).
  + Decision to obtain only the most important traits or to obtain all.
  + Default encoding to retrieve the content of the text is UTF-8.
  + Token key.

* **OUTPUT:**
  + Dataframe with the text, the traits discovered and the numeric probability.

An example is displayed next:

```{r get personality, eval=FALSE}
get_personality(url_api, text, language, all = FALSE, "UTF-8", token)
```

```{r get personality all, eval=FALSE}
get_personality(url_api, text, language, all = TRUE, "UTF-8", token)
```

## Communication style

The function **get_communic_style()** detects the communication purpose and style of the text. The style includes: 

* *self-revealing* (sharing one's experience and opinion). 

* *fact-oriented* (focusing on factual information, objective observations or statements). 

* *information-seeking* (posing questions). 

* *action-seeking* (aiming to trigger someone's action by giving recommendation, requests or advice).

Both the inputs and outputs are the same as in the case of personality traits. 

```{r get communication, eval=FALSE}
get_communic_style(url_api, text, language, all = FALSE, "UTF-8", token)
```

```{r get communication all, eval=FALSE}
get_communic_style(url_api, text, language, all = TRUE, "UTF-8", token)
```

## Emotions

The function **get_emotion()** allows the user to get the emotions expressed in a text. Retrieved emotions can be *anger*, *joy*, *love*, *sadness* or *surprise*. Supported languages here are "de" (German), "en" (English) and "es" (Spanish).

Both the inputs and outputs are the same as in previous cases, but here a set of maximum 512 texts is allowed.

```{r get emotions, eval=FALSE}
get_emotion(url_api, text, language, all = FALSE, "UTF-8", token)
```

```{r get emotions all, eval=FALSE}
get_emotion(url_api, text, language, all = TRUE, "UTF-8", token)
```

## Ekman's emotions

The function **get_ekman_emotion()** allows the user to get the Ekman's emotions expressed in a text. Retrieved emotions can be *anger*, *disgust*, *fear*, *joy*, *no-emotion*, *sadness* or *surprise*. Supported languages here are "de" (German), "en" (English) and "es" (Spanish).

Both the inputs and outputs are the same as in previous cases, but here a set of maximum 512 texts is allowed.

```{r get ekman emotions, eval=FALSE}
get_ekman_emotion(url_api, text, language, all = FALSE, "UTF-8", token)
```

```{r get ekman emotions all, eval=FALSE}
get_ekman_emotion(url_api, text, language, all = TRUE, "UTF-8", token)
```

## Sentiments

The function **get_sentiment()** allows the user to extract the sentiment expressed in a text. Retrieved sentiments can be *positive* or *negative*. Supported languages here are "de" (German), "en" (English) and "es" (Spanish).

Both the inputs and outputs are the same as in previous cases. Here a set of maximum 32 texts is allowed again.

```{r get sentiments, eval=FALSE}
get_sentiment(url_api, text, language, all = FALSE, "UTF-8", token)
```

```{r get sentiments all, eval=FALSE}
get_sentiment(url_api, text, language, all = TRUE, "UTF-8", token)
```

## Topic sentiments

The function **get_topic_sentiment()** detects topics and the topic category of each topic. Then, it analyzes the sentiment towards each of the topics mentioned. Supported languages here are "de" (German) and "en" (English).

Both the inputs and outputs are the same as in previous cases, excepting the parameter to show all probabilities, which is not used here. All the detected features will be displayed. The procedure goes here text by text.

```{r get topic sentiments, eval=FALSE}
get_topic_sentiment(url_api, text, language, "UTF-8", token)
```

## Language detection

The function **get_language_detection()** identifies the language of the input text. Any language can be detected. A set of maximum 64 texts is allowed. 

```{r get language, eval=FALSE}
get_language_detection(url_api, text, "UTF-8", token)
```

## Label texts

The function **label_text()** receives a given (Excel) file with a number of posts. Then, it assigns the psychological features that the user wants to. The texts are grouped into smaller subsets of texts of the same language, so the procedure will be faster than going text by text. 
In the interests of illustration, **symantoAPI** contains the file **excel_orig** with 50 toy texts. The first column is the text identifier. The second column contains the texts. The third column contains the language in which each text was written. 47 texts are written in English and 3 in German.

The function **label_text()** is going to be used with this data set.

Next the input and output of the function are detailed. 

* **INPUT:**
  + URL to the API.
  + Data frame with the texts to analyze. It must contain at least one column with the texts, named "Text" or "text". Another column with the language of the texts, named
"Language" or "language", is optional. Other complementary columns can be also added to be incorporated in the returned data frame.
  + User-defined psychological properties to evaluate. 
  + Default encoding to retrieve the content of the text is UTF-8.
  + Token key.
  + Texts in blocks: Number of texts in every block. 
  + If the language of the texts is not available, **get_language_detection()** will be used to assign a language to every text. For the few cases where this tool does not return a valid language, the assigned language is this parameter provided by the user. 
  In the case when the language is available, this parameter serves to replace those languages that are not allowed by the API. Default "en" (English).
  + Option to provide a flow of information.

* **OUTPUT:**
  + The input data frame filled with the outputs related to the psychological properties desired by the user.

**NOTE:** In order to load in R an external Excel file, the following code can be used:

```{r read excel, eval=FALSE}
library(readxl)
excel_orig <- read_excel("path_to_file/excel_file.xlsx")
```

The following example shows the results for a subset of psychological properties. In this case, only the colums related to personality traits and sentiments are filled. 

```{r sample label texts, eval=FALSE}
# The vector for all the actions is:
# user_actions <- c("personality", "communication", "emotions", "ekman_emotions", 
#                  "sentiments", "topic_sentiments", "language")
user_actions <- c("personality", "sentiments")  
excel_dest <- label_text(url_api, excel_orig, user_actions, "UTF-8", token, 32, "en", TRUE)
excel_dest
```

The output file can then be saved as an Excel file with:

```{r save excel, eval=FALSE}
library(openxlsx)
write.xlsx(excel_dest, "excel_dest.xlsx")
```

# Getting Help

For more information about the functions and data sets, the **help()** function and **?** help operator provide access to their documentation pages. For example:

```{r help, eval=FALSE}
?label_text
```

# Session information

```{r session info}
sessionInfo()
```